Untwisting the Tetrahelix (v0.8)

### 2017-03-01 doc/TetrahelixUntwist.tex
minor fixes

\documentclass[11pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage[affil-it]{authblk}
\usepackage[toc,page]{appendix}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

@@

\title{Untwisting the Tetrahelix (v0.7)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
called the Song-Kwon-Kim\cite{song2003spherical} or turret joint,
which allow many members to come to a single point.  Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

\begin{figure}[H] %float with two figures
  \label{fig:closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

@@

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in JavaScript and demonstrated by an interactive
design website\cite{readtetrahelix}, from which Figures \ref{fig:closeup} and
\ref{fig:series} are taken.

\begin{figure}[H] %float with two figures
  \label{fig:series}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices, BC Helix in Foreground.}
\end{figure}

%% Note: This is not rendering figure two for some reason.
In Figure
\ref{fig:series}, the closest helix is the BC helix, and the furthest
is the equitetrabeam.

@@

$\theta$ is approximately $131.81$ degrees.
The angle $\theta$ is the rotation of a \emph{each} tetrahedra, not the tetraheda along a rail.  In Figure \ref{closeup},
each tetrahedra has either a yellow, blue, or red outer edge or rail.
That is, a blue-rail tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the yellow tetrahedra.

@@

The apparent rotation of tetrhedra of the same outer-edge color ($V_3$ relative from $V_0$ for example,
in \eqref{graycoxeter}, is $3 \theta - 2\pi$.

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of deveoping a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 10
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

\section{Optimal Tetrahelices Have $z$-axis Evenly Spaced Vertices}

@@

\begin{theorem}
  \label{thm:cylinder}
  Any optimal tetrahelix has all vertices on a single cylinder.
\end{theorem}

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 \\
  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2    \numberthis  \label{railangle} \\
\end{align*}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(0) - r_{\rho}\sin(\rho/3+1\cdot\frac{2\pi}{3}))^2  +
    (r_{\rho}\cos(0) - r_{\rho}\cos(\rho/3 + 1\cdot\frac{2\pi}{3}))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin( \rho / 2)^2 \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.  
  \text{one-hop}  &= \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.    
\end{align*}

@@

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius is negative up until $\rho$ up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

@@

If you look down the axis of an optimal tetrahelix, it happens that only the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder or that must bear a human or to armor a conduit. The \emph{inradius} of
an optimal tetrahelix is a function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  \frac{r \sin{(\pi - \rho)}}{6}
\end{equation}
Which can be seen from the trigonometry of the projection diagram:

\begin{figure}[H]
  \label{projectiondiagram}
     \centering
     \includegraphics[width=0.3\textwidth]{figures/ProjectionDiagram.png}
     \caption{Projection Diagram (needs labels)}
\end{figure}

@@

This radius produces a two-hop rail length of
$\frac{2}{\sqrt{3}}$. The difference between this 
and $1$ of $\approx 15.47\% $.

\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $r = \sqrt{35}/4$,
which produces three length classes of $11/12, 12/12, 13/12$}

In Figure \ref{series}, the furthest tetrahelix is the optimal equitetrabeam.

>>>>

\documentclass[11pt]{article}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage[affil-it]{authblk}
\usepackage[toc,page]{appendix}
\usepackage{pifont}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{hyperref}

@@

\title{Untwisting the Tetrahelix (v0.8)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

The vertices of the tetrahedra lie upon
three helices about the central axis.  The
Tetrobot/Glussbot\cite{TetrobotBook} project uses the regularity of
this geometry to make a tentacle-like robot that can crawl like a slug
or mollusc.  The Tetrobot concept is to use mechanical members, called
actuators, which can change their length, connected by special joints,
such as the three-d printable Song-Kwon-Kim\cite{song2003spherical}, or the CMS joint\cite{HamlinSandersonCMS}
which allow many members to come to a single point.
Such machines can
follow purely regular mathematical models such as the Boerdijk–-Coxeter
helix or the Octet Truss\cite{richard1961synergetic}.

@@

\begin{figure}[H]
  \label{fig:spudboy}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/TetrahelixSeries.png}
     \caption{A Series of Tetrahelices}
\end{figure}

\begin{figure}[H]
  \label{fig:closeup}
  \centering
     \includegraphics[width=0.95\textwidth]{figures/BCHelixCloseUp.png}
     \caption{BC Helix Close-up}
\end{figure}

@@

BC helix does not rest on a plane in a simple way. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, we can
untwist a tetrahelix to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
deveoped here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix},
from which Figures \ref{fig:closeup} and
the figure below are taken.

In Figure \ref{fig:spudboy}, the closest helix is the BC helix, and the furthest
is the equitetrabeam.

@@

$\theta$ is approximately $131.81$ degrees.
The angle $\theta$ is the rotation of a \emph{each} tetrahedra, not the tetraheda along a rail.  In Figure \ref{fig:closeup},
each tetrahedra has either a yellow, blue, or red outer edge or rail.
That is, a blue-rail tetrahedron is rotated slightly more than a $1/3$ of a revolution to match the face of the yellow tetrahedra.

@@

The apparent rotation of tetrhedra of the same outer-edge color ($V_3$ relative from $V_0$ for example,
in \eqref{graycoxeter}, is $3 \theta - 2\pi$).

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
%%  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of deveoping a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 10
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

\section{Optimal Tetrahelices}

@@

\begin{theorem}
  \label{thm:cylinder}
  Any optimal tetrahelix with a rail angle less than $\pi$ has all vertices on a single cylinder.
\end{theorem}

@@

\begin{align*}
  1^2 &= d^2 + (2 r \sin{ \rho / 2})^2 \\
%%  1 &= d^2 + 4 r^2 (\sin{ \rho / 2})^2 \\
  d^2 &= 1 - 4 r^2 (\sin{ \rho / 2})^2    \numberthis  \label{railangle} \\
\end{align*}

@@

\begin{align*}
  \text{one-hop} &= dist(H_{general}(n,c,\rho,d_{\rho}),H_{general}(n,c+1,\rho,d_{\rho}),r_{\rho})  \\
%%  \text{one-hop} &= dist(H_{general}(0,0,\rho,d_{\rho}),H_{general}(0,1,\rho,d_{\rho},r_{\rho}))  \\  
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (r_{\rho}\sin(0) - r_{\rho}\sin(\rho/3+1\cdot\frac{2\pi}{3}))^2  +
%%    (r_{\rho}\cos(0) - r_{\rho}\cos(\rho/3 + 1\cdot\frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + (0  - r_{\rho}\sin(\rho/3 + \frac{2\pi}{3}))^2  +
%%    (r_{\rho} - r_{\rho}\cos(\rho/3 + \frac{2\pi}{3}))^2} \\
%%  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2\sin^2(\rho/3 + \frac{2\pi}{3})  +
%%    r_{\rho}^2(1 - \cos(\rho/3 + \frac{2\pi}{3}))^2} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.
  \text{one-hop}  &= \sqrt{\frac{d_{\rho}^2}{9} + r_{\rho}^2(\sin^2(\rho/3 + \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  \text{but: }  d_{\rho}^2 &= 1 - 4 r_{\rho}^2 (\sin( \rho / 2)^2 \text{so we substitute:}\\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.  
  \text{one-hop}  &= \sqrt{\frac{1}{9}  + r_{\rho}^2(-\frac{4 (\sin^2( \rho / 2))}{9} + \sin^2(\rho/3+ \frac{2\pi}{3})  + (1 - \cos(\rho/3 + \frac{2\pi}{3}))^2)} \\
  %% Formula above checks at sqrt(8/27), 1, 0, and rho_bc, h_bc, r_bc.    
\end{align*}

@@

By graph inspection using Mathematica, we see the partial derivative of this with respect to
radius $r_{\rho}$ is always negative.
Since the partial derivative of $\text{two-hop} - \text{one-hop}$ with respect to the
radius is negative up until $\rho_{bc}$ when it is $0$, then
we will opitmize the overall minimax distance by choosing the largest radius
up until one-hop $= 1$, the rail-edge length.

@@

\begin{align*}
  d_{opt}^2 &= 1 - 4 \frac{2}{\sqrt{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }}^2 (\sin{ \rho / 2})^2   \\
%%  d_{opt}^2 &= 1 - 4 \frac{4(\sin{ \rho / 2})^2}{\frac{9}{2} \cdot ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
  d_{opt}^2 &= 1 - \frac{16(\sin{ \rho / 2})^2}{9 ( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + \cos(\rho)+ 8 }    \\
%%  d_{opt}^2 &= 1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }   \\
    d_{opt} &= \sqrt{1 - \frac{16 \sin^2(\rho / 2) }{ \cos(\rho) + 9( \sqrt{3} \sin(\rho/3) + \cos(\rho/3)) + 8 }}    \numberthis  \label{dopt}  \\      
\end{align*}

@@

If you look down the axis of an optimal tetrahelix, it happens that only the one-hop edges
come closest to the axis. In other words, they define the radius of the incircle of the
projection, or the radius of a cylinder that would just fit inside the tetrahelix.
This is useful to know if the tetrahelix is used as a structure to house a pipe or
a ladder that must bear a human or to armor a conduit. The \emph{inradius} of
an optimal tetrahelix is a remarkably simple function of the radius $r$ and the rail angle $\rho$:
\begin{equation}
  \label{eq:inradius}
  \frac{r \sin{(\pi - \rho)}}{6}
\end{equation}
Which can be seen from the trigonometry of the projection diagram:

\begin{figure}[H]
  \label{projectiondiagram}
     \centering
     \includegraphics[width=0.8\textwidth]{figures/ProjectionDiagram.png}
     \caption{General One-hop Projection Diagram}
\end{figure}

@@

This radius produces a two-hop rail length of
$\frac{2}{\sqrt{3}}$\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $r = \sqrt{35}/4$,
which produces three length classes of $11/12, 12/12, 13/12$.}
. The difference between this 
and $1$ is $\approx 15.47\% $.

\begin{figure}[H]
  \label{fig:unfolded}
     \centering
     \includegraphics[width=0.6\textwidth]{figures/EquitetrabeamUnfolded.JPG}
     \caption{Projection Diagram (needs labels)}
\end{figure}

In Figure \ref{fig:spudboy}, the furthest tetrahelix is the optimal equitetrabeam.
Figure \ref{fig:unfolded} shows a representation of an unfolded equitetrabeam.
